---
title: "Data preprocessing"
author: "Ye Bi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# loading packages
```{r, message=F}
library(tidyverse)
library(lme4)
```

# 0. Grain-size phenotype data BLUE calculation
```{r}
wa_traits <- read.csv("~/Library/CloudStorage/OneDrive-VirginiaTech/Research/Codes/research/RiceUNLMetabolites/Trait/Waseem_trait/Rice_HNT_IMAGE_FINAL.csv")
wa_traits$treatment = gsub("Control", "control", wa_traits$treatment)
wa_traits$treatment = gsub("Stress", "stress", wa_traits$treatment)
majoraxis0 = wa_traits %>% group_by(treatment, NSFTV_ID) %>% summarise(C_MajorAxisLength_mean = mean(C_MajorAxisLength_mean))
majoraxis0 = majoraxis0[-which(majoraxis0$NSFTV_ID == 98), ]
minoraxis0 = wa_traits %>% group_by(treatment, NSFTV_ID) %>% summarise(C_MinorAxisLength_mean = mean(C_MinorAxisLength_mean))
perimeter0 = wa_traits %>% group_by(treatment, NSFTV_ID) %>% summarise(C_Perimeter_mean = mean(C_Perimeter_mean))
```

# 1. Metabolite data cleaning.
- Keep "NSFTV.Id"  "Treatment" "Ext.batch" "Run" and 73 Metabolites
- Remove NA, duplicated, no matched genotypes (NSFTV_ID). 
```{r, eval=F}
met22 <- read_excel("~/Library/CloudStorage/OneDrive-VirginiaTech/Research/Codes/research/RiceUNLMetabolites/Pheno/raw_data/221019_metabolite raw data.xlsx")
met22_info <- read_excel("~/Library/CloudStorage/OneDrive-VirginiaTech/Research/Codes/research/RiceUNLMetabolites/Pheno/raw_data/221019_metabolite raw data.xlsx", sheet = "Sample information")
colnames(met22_info)

all(met22_info$Sample == met22$Sample)

met = cbind.data.frame(NSFTV.Id = met22_info$NSFTV.Id, Treatment = met22_info$Treatment, Ext.batch = met22_info$Ext.batch, Run = met22_info$Run, met22[,-1]) # "NSFTV.Id"  "Treatment" "Ext.batch" "Run" and 73 Metabolites

colnames(met)[1] = "NSFTV_ID"
colnames(met)[3] = "Batch" 
met = met[!is.na(met$NSFTV_ID),] # one NA in NSFTV_ID
met$NSFTV_Id= paste0("NSFTV_",met$NSFTV_ID) 
met <- met %>% arrange(NSFTV_ID)


## Change met names into a1-h3.
name = c(paste0("a",1:10),paste0("b",1:10),paste0("c",1:10),paste0("d",1:10),paste0("e",1:10),paste0("f",1:10),paste0("g",1:10),paste0("h",1:3))
colnames(met)[5:77] = name
metL <- met %>% group_split(Treatment);names(metL) <- c("Control", "Stress")

#remove duplicated genotypes in met.stress (no dupicated genotypes in control group)
dd = duplicated(metL[[2]]$NSFTV_ID)
metL[[2]] = metL[[2]][!dd, ]

#remove the unused NSFTV_18. Since phenotype data in control group doesn't have it.
unuse = which(metL[[2]]$NSFTV_ID == "NSFTV_18")
metL[[2]] = metL[[2]][-unuse, ]
unuse1 = which(metL[[1]]$NSFTV_ID == "NSFTV_18")
metL[[1]] = metL[[1]][-unuse1, ]
save(metL, file="../metL_clean.RData")
```

# 2. Match genotype, phenotype, trait datasets.
```{r, eval = F}
setwd("~/Library/CloudStorage/OneDrive-VirginiaTech/Research/Codes/research/RiceUNLMetabolites/Trait/codes")
library(tidyverse)

###########################
# put three traits together
###########################
###Remove one genotype
majoraxis <- read.csv("../raw_data/grainlength.csv")
minoraxis <- read.csv("../raw_data/grainwidth.csv")
#NSFTV_98 just not show in minor axis and perimeter data set, so I decided to delete it.
minoraxis <- minoraxis[-which(minoraxis$NSFTV_ID == 98),]
perimeter <- read.csv("../raw_data/grainperimeter.csv")
perimeter <- perimeter[-which(perimeter$NSFTV_ID == 98),]

###Remove NA in NSFTV_ID
majoraxis=majoraxis[!is.na(majoraxis$NSFTV_ID),]
minoraxis=minoraxis[!is.na(minoraxis$NSFTV_ID),]
perimeter=perimeter[!is.na(perimeter$NSFTV_ID),]
table(is.na(majoraxis$NSFTV_ID))
table(is.na(minoraxis$NSFTV_ID))
table(is.na(perimeter$NSFTV_ID))

#Change names into more readable names.
majoraxis <- majoraxis[!is.na(majoraxis$NSFTV_ID),];colnames(majoraxis)[3]="MajorAxis"
minoraxis <- minoraxis[!is.na(minoraxis$NSFTV_ID),];colnames(minoraxis)[3]="MinorAxis"
perimeter <- perimeter[!is.na(perimeter$NSFTV_ID),];colnames(perimeter)[3]="Perimeter"

#Change NSFTV_ID from "number" into "NSFTV_ + number".
majoraxis$NSFTV_ID <- paste0("NSFTV_", majoraxis$NSFTV_ID)
minoraxis$NSFTV_ID <- paste0("NSFTV_", minoraxis$NSFTV_ID)
perimeter$NSFTV_ID <- paste0("NSFTV_", perimeter$NSFTV_ID)

#Split control and stress and save in one list.
majoraxisL <- majoraxis %>% group_split(treatment)
minoraxisL <- minoraxis %>% group_split(treatment)
perimeterL <- minoraxis %>% group_split(treatment)

#Merge three traits and save as csv file.
three_traits <- merge(merge(majoraxis,minoraxis),perimeter)
write.csv(three_traits, "../three_traits.csv", row.names = F, quote = F)

#Split traits into control and stress groups, then save.
three_traitsL <- three_traits %>% group_split(treatment)
three_traits_control <- three_traitsL[[1]];colnames(three_traits_control)[1]="Treatment"
three_traits_stress <- three_traitsL[[2]];colnames(three_traits_stress)[1]="Treatment"
write.csv(three_traits_control,"../three_traits_control.csv", row.names = F, quote = F)
write.csv(three_traits_stress, "../three_traits_stress.csv", row.names = F, quote = F)


######################################################
##Match phenotyped, genotyped, metabolited data set.
######################################################
# 1. Loading data
## 1.1 Read geno data.
geno_700K <- readRDS("~/Library/CloudStorage/OneDrive-VirginiaTech/Research/Codes/research/RiceUNLMetabolites/Geno/geno_final_700K.rds")
map <- readRDS("~/Library/CloudStorage/OneDrive-VirginiaTech/Research/Codes/research/RiceUNLMetabolites/Geno/mapinfo_final_700k.rds")
SNP_names <- map$V2
colnames(geno_700K) = SNP_names
geno_id <- rownames(geno_700K)

path.met = "../../Pheno"
path.trait = "../../Trait"

## 1.2 Read metabolite data
###this is list after cleaning duplicated and unuseful metabolites.
load(file=file.path(path.met, "metL_clean.RData"))  #metL

## 1.3 Phenotype (traits) data
trait.con <- read.csv(file=file.path(path.trait, "three_traits_control.csv"))
trait.trt <- read.csv(file=file.path(path.trait, "three_traits_stress.csv"))
three_traitsL <- list(trait.con, trait.trt)

#function to find out overlap in control and stress
id_func <- function(dfL){
  length(unique(dfL[[1]]$NSFTV_ID)) #234 #so group1 is control group.
  length(unique(dfL[[2]]$NSFTV_ID)) #229
  ix <- dfL[[1]]$NSFTV_ID %in% dfL[[2]]$NSFTV_ID #192 42
  ids_c_t = dfL[[1]]$NSFTV_ID[ix]
  ids_c_others = dfL[[1]]$NSFTV_ID[!ix]
  table(dfL[[2]]$NSFTV_ID %in% dfL[[1]]$NSFTV_ID) #192 38
  ix1 <- dfL[[2]]$NSFTV_ID %in% dfL[[1]]$NSFTV_ID
  ids_t_others <- dfL[[2]]$NSFTV_ID[!ix1]
  ids_c_t1 <- dfL[[2]]$NSFTV_ID[ix1]
  ll <- list(ids_c_t=ids_c_t,ids_t_c=ids_c_t1, ids_c_others=ids_c_others, ids_t_others=ids_t_others)
  return(ll)
}

three_traits_ll <- id_func(three_traitsL)
met_ll <- id_func(metL)


#function to find out overlap with genotype data.
id_match <- function(dfL){
        ix2 <- dfL[[1]]$NSFTV_ID %in% geno_id
        ids_c_g <- dfL[[1]]$NSFTV_ID[ix2]
        ix3 <- dfL[[2]]$NSFTV_ID %in% geno_id
        ids_t_g <- dfL[[2]]$NSFTV_ID [ix3]
        ll <- list(ids_c_g = ids_c_g, ids_t_g = ids_t_g)
        return(ll)
}


geno_trait<- id_match(three_traitsL) #overlap between metabolite and geno data.
met_geno_trait <- id_match(metL) #since all metabolites are phenotyped, so this is also overlap among met&geno&trait.

## 1.3 clean met and trait dataset.
met.con <- as.data.frame(metL[[1]])
met.trt <- as.data.frame(metL[[2]])

met.con <- met.con[met.con$NSFTV_ID %in% met_geno_trait$ids_c_g, ] #192*77 run batch
met.trt <- met.trt[met.trt$NSFTV_ID %in% met_geno_trait$ids_t_g, ] #188*77 run batch
trait.con <- trait.con[trait.con$NSFTV_ID %in% met_geno_trait$ids_c_g, ] #192*5
trait.trt <- trait.trt[trait.trt$NSFTV_ID %in% met_geno_trait$ids_t_g, ] #188*5


##change all "control" to "Control", "stress" to "Stress"
trait.con$Treatment = "Control"
trait.trt$Treatment = "Stress"
#save all the matched met and trait files.
save(met.con, file=file.path(path.met, "met.con.match.Rdata"))
save(met.trt, file=file.path(path.met, "met.trt.match.Rdata"))
save(trait.con, file = file.path(path.trait, "trait.con.Rdata"))
save(trait.trt, file = file.path(path.trait, "trait.trt.Rdata"))
```


# 3. Adjust metabolites
- Relative metabolite abundance was corrected for run and experimental batch effects by treating them as random, separately for the control and HNT conditions.
```{r, eval = F}
 # control group
met.con = read.csv("../met.control.noNA.csv")
met.con = met.con %>% arrange(NSFTV_ID)
met.con <- met.con %>% mutate(across(c(Run, Batch, NSFTV_ID), as.factor))

# treated group
met.trt = read.csv("../met.stress.noNA.csv")
met.trt = met.trt %>% arrange(NSFTV_ID)
met.trt <- met.trt %>% mutate(across(c(Run, Batch, NSFTV_ID), as.factor))

correct_met_func <- function(x, n_met = 73, treatment){
    # x = met.con
    # n_met = 73
    # treatment="control"
    rr = matrix(NA, nrow = nrow(x), ncol = n_met); colnames(rr) = colnames(x)[-c(1:4)]
    for (i in colnames(rr)){
      # i="a1"
      formu = as.formula(paste0(i, " ~ 1+(1|Run) + (1|Batch)"))
      print(formu)
      # fit <- lm(formula=formu, data=x)
      # summary(fit)
      fit1 = lmer(formula = formu, data = x)
      summary(fit1)
      rrr = residuals(fit1)
      rr[,i] = rrr
    }
    met.rr = cbind.data.frame(x[,1:2], rr)
    # treatment = "con"
    write.csv(met.rr, paste0("../met.rr.",treatment, ".csv"), row.names = F, quote = F)
}
# correct_met_func(x=met.comb, n_met=73, treatment="comb")

# 
correct_met_func(met.con, n_met = 73, treatment="con")
correct_met_func(met.trt, n_met = 73, treatment="trt")
```

# 4. Genotype data cleaning.
```{r, eval = F}

## Read into met.rr dataset to get NSFTV_ID in treatments.
met_rr.con <- read_csv("~/Library/CloudStorage/OneDrive-VirginiaTech/Research/Codes/research/RiceUNLMetabolites/Pheno/met.rr.con.csv")
met_rr.trt <- read_csv("~/Library/CloudStorage/OneDrive-VirginiaTech/Research/Codes/research/RiceUNLMetabolites/Pheno/met.rr.trt.csv")
id_con <- met_rr.con$NSFTV_ID
id_trt <- met_rr.trt$NSFTV_ID

## Read into raw genotype data.
geno_700k <- readRDS("geno_final_700K.rds") #SNPs
map <- readRDS("mapinfo_final_700k.rds") #map
SNP_names <- map$V2
colnames(geno_700k) = SNP_names

####################################
#### Control group #################
####################################
geno_con <- geno_700k[id_con,]
#MAF check
p <- colMeans(geno_con)/2
maf <- ifelse(p> 0.5, 1-p, p)
maf.index <- which(maf < 0.05) 
geno_con <- geno_con[,-maf.index] # 192 385118
save(geno_con, file="geno.rr.con.RData")


####################################
#### Stress group #################
####################################
geno_trt <- geno_700k[id_trt,]
#MAF check
p <- colMeans(geno_trt)/2
maf <- ifelse(p> 0.5, 1-p, p)
maf.index <- which(maf < 0.05) 
geno_trt <- geno_trt[,-maf.index] # 188 389854
save(geno_trt, file="geno.rr.trt.RData")
```

# 5. Subpopulation summary
```{r, eval = F}
panel <- read.delim("~/Library/CloudStorage/OneDrive-VirginiaTech/Research/Codes/research/RiceUNLMetabolites/Geno/raw_data/panel_info_clean.txt")
panel <- panel %>% filter(Sample.set == "RDP1")
panel$NSFTV_ID <- gsub("NSFTV","NSFTV_", panel$Other.accession.ID)

load("~/Library/CloudStorage/OneDrive-VirginiaTech/Research/Codes/research/RiceUNLMetabolites/Trait/trait.trt.Rdata")
load("~/Library/CloudStorage/OneDrive-VirginiaTech/Research/Codes/research/RiceUNLMetabolites/Trait/trait.con.Rdata")
trait <- rbind(trait.con, trait.trt)
lines <- unique(trait$NSFTV_ID)

our <- panel[panel$NSFTV_ID %in% lines, ]

#percentage of subpopulation
round(table(our$fastStructure.subpopulation.call)/219*100, digit=2)
```

