---
title: "GBLUP_MBLUP_GMBLUP_Subpop"
author: "Ye Bi"
date: "2023-01-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/OneDrive - Virginia Tech/Research/Codes/research/RiceUNLMetabolites/Prediction/all_mets/HeatmapPCA")
path.met = "../../../Pheno"
path.trait = "../../../Trait"
path.geno = "../../../Geno"
path.output = "./outputs"
```

### loading packages
```{r, message=F}
# library(rrBLUP)
library(BGLR)
library(devtools)
library(ggbiplot)
library(pheatmap)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(ggfortify)#draw pca plots
library(data.table)
```

```{r}
met_names <- read.delim("~/Library/CloudStorage/OneDrive-VirginiaTech/Research/Codes/research/RiceUNLMetabolites/Pheno/raw_data/met_names.txt", header=FALSE)
names <- met_names$V1
```

## Read met data.
```{r}
met.con <- read.csv(file=file.path(path.met, "met.rr.con.csv")) #192*75 n
colnames(met.con)[-c(1:2)] = names
met.trt <- read.csv(file=file.path(path.met, "met.rr.trt.csv")) #188*75
colnames(met.trt)[-c(1:2)] = names
```

## Read traits data
```{r}
load(file=file.path(path.trait, "trait.con.Rdata"))
load(file=file.path(path.trait, "trait.trt.Rdata"))
# trait.con$NSFTV_ID <- as.factor(trait.con$NSFTV_ID)
# trait.trt$NSFTV_ID <- as.factor(trait.trt$NSFTV_ID)
```

## Read geno data.

```{r}
load(file=file.path(path.geno, "geno.rr.con.RData")) # 192 385118
load(file=file.path(path.geno, "geno.rr.trt.RData")) # 188 389854
```

## G matrix and M matrix
### control
```{r, eval = F}
# G matrix
Gcs.con = scale(geno_con, center = T, scale = T)
Gchnt.con = tcrossprod(Gcs.con)/ncol(Gcs.con)

# M matrix
rownames(met.con) = met.con$NSFTV_ID
Mcs.con = scale(met.con[,-c(1:2)]) 
Mchnt.con = tcrossprod(Mcs)/ncol(Mcs.con)# for Bayesian MBLUP
```

### HNT
```{r, eval = F}
# G matrix
Gcs.trt = scale(geno_trt, center = T, scale = T)
Gchnt.trt = tcrossprod(Gcs.trt)/ncol(Gcs.trt)

# M matrix
rownames(met.trt) = met.trt$NSFTV_ID
Mcs.trt = scale(met.trt[,-c(1:2)]) 
Mchnt.trt = tcrossprod(Mcs.trt)/ncol(Mcs.trt)# for Bayesian MBLUP
```

## sub info
```{r}
#read into subpop info
panel_info_clean <- read.delim("~/Library/CloudStorage/OneDrive-VirginiaTech/Research/Codes/research/RiceUNLMetabolites/Geno/raw_data/panel_info_clean.txt")
subpop <- panel_info_clean %>% filter(Sample.set == "RDP1")
subpop$Other.accession.ID = gsub("NSFTV", "NSFTV_", subpop$Other.accession.ID)

table(subpop$Other.accession.ID %in% trait.con$NSFTV_ID)
table(subpop$Other.accession.ID %in% trait.trt$NSFTV_ID) #188
table(trait.con$NSFTV_ID %in% subpop$Other.accession.ID) #all in suspop txt file.
table(trait.trt$NSFTV_ID %in% subpop$Other.accession.ID)  #all
```

## Function for combine geno and subpop in one dataframe.
```{r, eval = F}
pca_sub_func=function(trait){
            pop.df = subpop[subpop$Other.accession.ID %in% trait$NSFTV_ID,]
            pop.df = data.frame(NSFTV_ID = pop.df$Other.accession.ID,
                                Subpopu = pop.df$fastStructure.subpopulation.call)
            pop.df = pop.df %>% arrange(NSFTV_ID)
}

pop.df.con = pca_sub_func(trait.con)
pop.df.trt = pca_sub_func(trait.trt)
save(pop.df.con, pop.df.trt, file=file.path(path.output, "pop.df.rr.RData"))

#check if ids are in order
table(pop.df.con$NSFTV_ID == trait.con$NSFTV_ID)
table(pop.df.trt$NSFTV_ID == trait.trt$NSFTV_ID)
```



### Conventional H2
```{r}
library(lme4)
library(readxl)
Grain_Size <- read_excel("~/Library/CloudStorage/OneDrive-VirginiaTech/Research/Codes/research/RiceUNLMetabolites/Trait/Waseem_trait/Grain_Size_HNT_JS.xlsx", 
    sheet = "AllSeeds")
colnames(Grain_Size)[2] = "GSOR.ID"
Grain_Size$GSOR.ID = paste0("GSOR", Grain_Size$GSOR.ID)
Grain_size.con = Grain_Size[Grain_Size$Treatment == "C", ]
Grain_size.trt = Grain_Size[Grain_Size$Treatment == "S", ]

Grain_size.rep.con = Grain_size.con%>%group_by(GSOR.ID) %>% summarise(nn = length(unique(PlantNo.)))
rep.con = mean(Grain_size.rep.con$nn)
Grain_size.rep.trt = Grain_size.trt%>%group_by(GSOR.ID) %>% summarise(nn = length(unique(PlantNo.)))
rep.trt = mean(Grain_size.rep.trt$nn)

##### Control
Grain_size.con$GSOR.ID = as.factor(Grain_size.con$GSOR.ID)
Grain_size.con$PlantNo. = as.factor(Grain_size.con$PlantNo.)
traits = c("MajorAxisLength", "MinorAxisLength", "Perimeter")
for (i in 1:3){
formu = as.formula(paste0(traits[i],"~ 1 + (1|GSOR.ID) + PlantNo."))
fit.lme = lmer(formula = formu, data = Grain_size.con)
summary(fit.lme)
temp = as.data.frame(VarCorr(fit.lme))

H2 = temp[1,4]/(temp[1,4]+temp[2,4]/rep.con)
print(round(H2, 2))
}

##### Stress
Grain_size.trt$GSOR.ID = as.factor(Grain_size.trt$GSOR.ID)
Grain_size.trt$PlantNo. = as.factor(Grain_size.trt$PlantNo.)
traits = c("MajorAxisLength", "MinorAxisLength", "Perimeter")
for (i in 1:3){
formu = as.formula(paste0(traits[i],"~ 1 + (1|GSOR.ID) + PlantNo."))
fit.lme = lmer(formula = formu, data = Grain_size.trt)
summary(fit.lme)
temp = as.data.frame(VarCorr(fit.lme))

H2 = temp[1,4]/(temp[1,4]+temp[2,4]/rep.trt)
print(round(H2, 2))
}
```


##### Drawing PCA plots for subpopulation information
### loading packages
```{r, message=F}
# library(rrBLUP)
# library(BGLR)
library(devtools)
library(ggbiplot)
library(pheatmap)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(ggfortify)#draw pca plots
library(data.table)
```

```{r}
met_names <- read.delim("~/Library/CloudStorage/OneDrive-VirginiaTech/Research/Codes/research/RiceUNLMetabolites/Pheno/raw_data/met_names.txt", header=FALSE)
names <- met_names$V1
```

## Read met data.
```{r}
met.con <- read.csv(file=file.path(path.met, "met.rr.con.csv")) #192*75
colnames(met.con)[-c(1:2)] = names
met.trt <- read.csv(file=file.path(path.met, "met.rr.trt.csv")) #188*75
colnames(met.trt)[-c(1:2)] = names
```

## Read traits data
```{r}
load(file=file.path(path.trait, "trait.con.Rdata"))
load(file=file.path(path.trait, "trait.trt.Rdata"))
trait.con$NSFTV_ID <- as.factor(trait.con$NSFTV_ID)
trait.trt$NSFTV_ID <- as.factor(trait.trt$NSFTV_ID)
```

## Read geno data.

```{r}
load(file=file.path(path.geno, "geno.rr.con.RData")) # 192 385118
load(file=file.path(path.geno, "geno.rr.trt.RData")) # 188 389854
```
```{r}
## Read into raw genotype data.
geno_700k <- readRDS(file = file.path(path.geno, "/raw_data/geno_final_700K.rds")) #SNPs
map <- readRDS(file=file.path(path.geno, "/raw_data/mapinfo_final_700k.rds")) #map
SNP_names <- map$V2
colnames(geno_700k) = SNP_names

id_comb = unique(c(rownames(geno_con), rownames(geno_trt))) #219

geno_comb <- geno_700k[id_comb,] # 219 411066
#MAF check
p <- colMeans(geno_comb)/2
maf <- ifelse(p> 0.5, 1-p, p)
maf.index <- which(maf < 0.05) 
geno_comb <- geno_comb[,-maf.index] # 219 388099
# save(geno_comb, file="./PCA_subpop_outputs/geno.rr.comb.RData")
```

## G matrix and M matrix
```{r}
# G matrix
Gcs = scale(geno_comb, center = T, scale = T)
Gchnt = tcrossprod(Gcs)/ncol(Gcs)
```

## Control
```{r}
# M matrix
rownames(met.con) = met.con$NSFTV_ID
Mcs.con = scale(met.con[,-c(1:2)]) 
Mchnt.con = tcrossprod(Mcs.con)/ncol(Mcs.con)# for Bayesian MBLUP
```

## HNT
```{r}
# M matrix
rownames(met.trt) = met.trt$NSFTV_ID
Mcs.trt = scale(met.trt[,-c(1:2)]) 
Mchnt.trt = tcrossprod(Mcs.trt)/ncol(Mcs.trt)# for Bayesian MBLUP
```

## sub info
```{r}
#read into subpop info
panel_info_clean <- read.delim("~/Library/CloudStorage/OneDrive-VirginiaTech/Research/Codes/research/RiceUNLMetabolites/Geno/raw_data/panel_info_clean.txt")
subpop <- panel_info_clean %>% filter(Sample.set == "RDP1")
subpop$Other.accession.ID = gsub("NSFTV", "NSFTV_", subpop$Other.accession.ID)

table(subpop$Other.accession.ID %in% rownames(geno_comb))
table(rownames(geno_comb) %in% subpop$Other.accession.ID) #all in suspop txt file.
```

## Function for combine geno and subpop in one dataframe.
```{r}
pca_sub_func=function(geno, Gchnt){
            pop.df = subpop[subpop$Other.accession.ID %in% rownames(geno),]
            pop.df = data.frame(NSFTV_ID = pop.df$Other.accession.ID,
                                  Subpopu = pop.df$fastStructure.subpopulation.call)

            G.df = data.frame(NSFTV_ID = rownames(Gchnt), Gchnt)
            G.suspop = merge(G.df, pop.df)
            return(G.suspop)
            }
```

```{r}
G.suspop = pca_sub_func(geno = geno_comb, Gchnt = Gchnt)
table(G.suspop$Subpopu)
```
#control
```{r, echo = TRUE, eval=TRUE}
G.suspop$Subpopulation = G.suspop$Subpopu

cc.pca = prcomp(G.suspop[,-c(1,221,222)], scale. = T)

autoplot(cc.pca, data = G.suspop, colour = "Subpopulation") + 
  theme_bw()
dev.print(pdf, file = file.path(path.output, "./pca_geno_comb.pdf"), height = 6, width = 8)
```

#stress
```{r}
G.trt.suspop = pca_sub_func(geno = geno_trt, Gchnt = Gchnt.trt)
table(G.trt.suspop$Subpopu)
```

```{r, echo = TRUE, eval=TRUE}
cc.pca = prcomp(G.trt.suspop[,-c(1,190)], scale. = T)
autoplot(cc.pca, data = G.trt.suspop, colour = "Subpopu") +
  theme_bw()
dev.print(pdf, file = file.path(path.output, "./pca_geno_stress.pdf"), height = 6, width = 8)
```

#control--M
```{r}

M.con.suspop = pca_sub_func(geno = met.con[,-c(1:2)], Gchnt = Mchnt.con)
table(M.con.suspop$Subpopu)
```

### follow reviewers' suggestions to compare top 5 ids, but not include in manuscript
```{r}
table(trait.con$NSFTV_ID %in% trait.trt$NSFTV_ID)
table(trait.trt$NSFTV_ID %in% trait.con$NSFTV_ID)

trait.con.shared = trait.con[trait.con$NSFTV_ID %in% trait.trt$NSFTV_ID, ] #161 5
trait.trt.shared = trait.trt[trait.trt$NSFTV_ID %in% trait.con$NSFTV_ID, ] #161 5


all(trait.con.shared$NSFTV_ID == trait.trt.shared$NSFTV_ID)

trait.diff = data.frame(NSFTV_ID = trait.con.shared$NSFTV_ID, abs((trait.con.shared[,-c(1:2)] - trait.trt.shared[,-c(1:2)])/trait.trt.shared[,-c(1:2)] * 100))

summary(trait.diff)

major_order = trait.diff[order(trait.diff$MajorAxis, decreasing = T), ]
major_top_id = major_order$NSFTV_ID[1:5]
  
minor_order = trait.diff[order(trait.diff$MinorAxis, decreasing = T), ]
minor_top_id = minor_order$NSFTV_ID[1:5]

perimeter_order = trait.diff[order(trait.diff$Perimeter, decreasing = T), ]
perimeter_top_id = perimeter_order$NSFTV_ID[1:5]
```

##MajorAxis
```{r}
met.con.top5.major = met.con[met.con$NSFTV_ID %in% major_top_id, -c(1:2)]
met.trt.top5.major = met.trt[met.trt$NSFTV_ID %in% major_top_id, -c(1:2)]

for( i in 1:5){
  cat("Corr is: ", cor(as.numeric(met.con.top5.major[i, ]), as.numeric(met.trt.top5.major[i, ])), "\n")
  t_test = t.test(as.numeric(met.con.top5.major[i, ]),
                  as.numeric(met.trt.top5.major[i, ]),
                  paired = TRUE)
  cat("t-test p-value: ", t_test$p.value, "\n")
}
```

##MinorAxis
```{r}
met.con.top5.minor = met.con[met.con$NSFTV_ID %in% minor_top_id, -c(1:2)]
met.trt.top5.minor = met.trt[met.trt$NSFTV_ID %in% minor_top_id, -c(1:2)]

for( i in 1:5){
  cat("Corr is: ", cor(as.numeric(met.con.top5.minor[i, ]), as.numeric(met.trt.top5.minor[i, ])), "\n")
  t_test = t.test(as.numeric(met.con.top5.minor[i, ]),
                  as.numeric(met.trt.top5.minor[i, ]),
                  paired = TRUE)
  cat("t-test p-value: ", t_test$p.value, "\n")
}
```

##Perimeter
```{r}
met.con.top5.perimeter = met.con[met.con$NSFTV_ID %in% perimeter_top_id, -c(1:2)]
met.trt.top5.perimeter = met.trt[met.trt$NSFTV_ID %in% perimeter_top_id, -c(1:2)]

for( i in 1:5){
  cat("Corr is: ", cor(as.numeric(met.con.top5.perimeter[i, ]), as.numeric(met.trt.top5.perimeter[i, ])), "\n")
  t_test = t.test(as.numeric(met.con.top5.perimeter[i, ]),
                  as.numeric(met.trt.top5.perimeter[i, ]),
                  paired = TRUE)
  cat("t-test p-value: ", t_test$p.value, "\n")
}
```


