---
title: "Section5 Genomic Prediction"
author: "Ye Bi"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/OneDrive - Virginia Tech/Research/Codes/research/RiceUNLMetabolites/Prediction/Prediction")
path.met = "../../Pheno"
path.trait = "../../Trait"
path.output = "./outputs"
path.geno = "../../Geno"
```

### Loading packages
```{r, message=F}
library(rrBLUP)
library(BGLR)
library(ggplot2)
library(ggpubr)
library(tidyverse)
```

# 1. loading data
## 1.1 Read met data
```{r, echo = T, eval = T}
met.con <- read.csv(file=file.path(path.met, "met.rr.con.csv")) #192*75
met.trt <- read.csv(file=file.path(path.met, "met.rr.trt.csv")) #188*75
```

## 1.2 Trait data: Major axis, Minor axis and  Perimeter data
```{r, echo = TRUE, eval = T}
load(file = file.path(path.trait, "trait.con.Rdata"))
load(file = file.path(path.trait, "trait.trt.Rdata"))
trait.con$NSFTV_ID <- as.factor(trait.con$NSFTV_ID)
trait.trt$NSFTV_ID <- as.factor(trait.trt$NSFTV_ID)
```
## 1.3 Read SNP data
```{r, eval = F}
load(file=file.path(path.geno, "geno.rr.con.RData")) # 192 385118
load(file=file.path(path.geno, "geno.rr.trt.RData")) # 188 389854
```


# 2. Prediction for control
## 2.1 GBLUP and GMBLUP
```{r, echo=T, eval=F}
# Scale traits (Normalize)
y0 = scale(trait.con[,-c(1:2)], center = T, scale = T)

# G matrix
Gcs = scale(geno_con, center = T, scale = T)
Gchnt = tcrossprod(Gcs)/ncol(Gcs)
EVD_Gchnt <- eigen(Gchnt)

# M matrix
# M matrix 

rownames(met.con) = met.con$NSFTV_ID
Mcs = scale(met.con[,-c(1:2)]) 
Mchnt = tcrossprod(Mcs) # for Bayesian MBLUP
Mchnt2 = Mchnt/mean(diag(Mchnt))
EVD_Mchnt <- eigen(Mchnt2)

#GBLUP
ETA1 <- list(
    G = list(V=EVD_Gchnt$vectors, d=EVD_Gchnt$values, model='RKHS')
   )

#MBLUP
ETA2 <- list(
        MET = list(V=EVD_Mchnt$vectors, d=EVD_Mchnt$values, model='RKHS')
   )

#GMBLUP
ETA3 <- list(
    G = list(V=EVD_Gchnt$vectors, d=EVD_Gchnt$values, model='RKHS'),
    MET = list(V=EVD_Mchnt$vectors, d=EVD_Mchnt$values, model='RKHS')
   )

# set cross-validation parameters
ntst  <-  38 # test 192*0.2
ntrn  <- 192-38 # reference 192*0.8
nCV <- 100 # times of CV

traits <- c("MajorAxis", "MinorAxis", "Perimeter")
# predictive correlation
corR1 <- matrix(0, ncol = ncol(y0), nrow = nCV) #OLS
  colnames(corR1) = traits
corR2 <- matrix(0, ncol = ncol(y0), nrow = nCV) #MBLUP
  colnames(corR2) = traits
corR3 <- matrix(0, ncol = ncol(y0), nrow = nCV) #BayesC
  colnames(corR3) = traits

# CV
for (i in 1:nCV) {
  # i=1
  # random-sampling to decide testing & reference accessions
  set.seed(100 + i)
  index <- sample(1:nrow(y0), size=ntst) # random sampling
  y <- y0
  y[index,] <- NA
  for (j in 1:ncol(y0)){
  cat("Now running nCV = ", i,"trait = ", j, "\n")
  ySingle <- y[,j]
  
  #GBLUP
  fit1 <- BGLR(y=ySingle, ETA=ETA1, nIter = 30000,
             burnIn = 10000, thin = 5, verbose = F, saveAt = 'GBLUP_')
  pred1 = fit1$yHat
  corR1[i,j] <- cor(pred1[index], y0[index,j])
  
  
  #GMBLUP
  fit3 <- BGLR(y=ySingle, ETA=ETA3, nIter = 30000,
             burnIn = 10000, thin = 5, verbose = F, saveAt = 'GMBLUP_')
  pred3 = fit3$yHat
  corR3[i,j] <- cor(pred3[index], y0[index,j])
  }}

save(corR1, corR3, file=file.path(path.output, "control_GBLUP_GMBLUP.rda"))
```

### 2.1.1 Calculate colmean
```{r}
load(file=file.path(path.output, "control_GBLUP_GMBLUP.rda"))
colnames(corR1) = c("Major_Axis", "Minor_Axis", "Perimeter")
round(colMeans(corR1), digits = 4) #GBLUP

# colnames(corR2) = c("Major_Axis", "Minor_Axis", "Perimeter")
# round(colMeans(corR2), digits = 4)#MBLUP

colnames(corR3) = c("Major_Axis", "Minor_Axis", "Perimeter")
round(colMeans(corR3), digits = 4) #GMBLUP
```

### 2.1.2 Heritability
```{r, echo=TRUE, eval = FALSE}
# Scale traits (Normalize)
y0 = scale(trait.con[,-c(1:2)], center = T, scale = T)

# G matrix
Gcs = scale(geno_con, center = T, scale = T)
Gchnt = tcrossprod(Gcs)/ncol(Gcs)
EVD_Gchnt <- eigen(Gchnt)

# M matrix
# M matrix 
Mcs = scale(met.con[,-c(1:2)]) #since met.rr was scaled when I did BLUP correction, so here no necessary to double normalization.
Mchnt = tcrossprod(Mcs) # for Bayesian MBLUP
Mchnt2 = Mchnt/mean(diag(Mchnt))
EVD_Mchnt <- eigen(Mchnt2)

#GBLUP
ETA1 <- list(
    G = list(V=EVD_Gchnt$vectors, d=EVD_Gchnt$values, model='RKHS')
   )

#MBLUP
ETA2 <- list(
        MET = list(V=EVD_Mchnt$vectors, d=EVD_Mchnt$values, model='RKHS')
   )

#GMBLUP
ETA3 <- list(
    G = list(V=EVD_Gchnt$vectors, d=EVD_Gchnt$values, model='RKHS'),
    MET = list(V=EVD_Mchnt$vectors, d=EVD_Mchnt$values, model='RKHS')
   )



for (j in 1:3){
  cat("Now running trait: ", colnames(y0)[j], "\n")
  set.seed(100+j)
  ySingle <- y0[,j]
  #GBLUP
  fit1 <- BGLR(y=ySingle, ETA=ETA1, nIter = 30000, 
             burnIn = 10000, thin = 5, verbose = F)
  
  varE = fit1$varE
  varM = fit1$ETA$G$varU
  h2M = varM/(varM+varE)
  cat("fit1 heritability for genomics is: ", round(h2M, 4), "\n")
  
  # MBLUP
  fit2 <- BGLR(y=ySingle, ETA=ETA2, nIter = 30000, 
             burnIn = 10000, thin = 5, verbose = F)
  varE = fit2$varE
  varM = fit2$ETA$MET$varU
  h2M = varM/(varM+varE)
  cat("fit2 heritability for metabolites is: ", round(h2M,4), "\n")
  
  #GMBLUP
  fit3 <- BGLR(y=ySingle, ETA=ETA3, nIter = 30000, 
             burnIn = 10000, thin = 5, verbose = F)
  varE = fit3$varE
  varM = fit3$ETA$MET$varU
  varG = fit3$ETA$G$varU
  h2G = varG/(varG+varM+varE)
  cat("fit3 heritability for genomics is: ", round(h2G,4), "\n")
  h2M= varM/(varG+varE+varM)
  cat("fit3 heritability for metabolites is: ", round(h2M,4), "\n")
  }
```
Now running trait:  MajorAxis 
fit1 heritability for genomics is:  0.7836 
fit2 heritability for metabolites is:  0.3254 
fit3 heritability for genomics is:  0.6988 
fit3 heritability for metabolites is:  0.0893 
Now running trait:  MinorAxis 
fit1 heritability for genomics is:  0.7545 
fit2 heritability for metabolites is:  0.4351 
fit3 heritability for genomics is:  0.6312 
fit3 heritability for metabolites is:  0.1336 
Now running trait:  Perimeter 
fit1 heritability for genomics is:  0.7533 
fit2 heritability for metabolites is:  0.3307 
fit3 heritability for genomics is:  0.6257 
fit3 heritability for metabolites is:  0.1167 


# 3. Prediction for stress
## 3.1  GBLUP and GMBLUP
```{r, echo=T, eval=F}
# Scale traits (Normalize)
y0 = scale(trait.trt[,-c(1:2)], center = T, scale = T)

# G matrix
Gcs = scale(geno_trt, center = T, scale = T)
Gchnt = tcrossprod(Gcs)/ncol(Gcs)
EVD_Gchnt <- eigen(Gchnt)

# M matrix
# M matrix 
Mcs = scale(met.trt[,-c(1:2)]) 
Mchnt = tcrossprod(Mcs) # for Bayesian MBLUP
Mchnt2 = Mchnt/mean(diag(Mchnt))
EVD_Mchnt <- eigen(Mchnt2)

#GBLUP
ETA1 <- list(
    G = list(V=EVD_Gchnt$vectors, d=EVD_Gchnt$values, model='RKHS')
   )

#MBLUP
ETA2 <- list(
        MET = list(V=EVD_Mchnt$vectors, d=EVD_Mchnt$values, model='RKHS')
   )

#GMBLUP
ETA3 <- list(
    G = list(V=EVD_Gchnt$vectors, d=EVD_Gchnt$values, model='RKHS'),
    MET = list(V=EVD_Mchnt$vectors, d=EVD_Mchnt$values, model='RKHS')
   )

# set cross-validation parameters
ntst  <-  38 # test 188*0.2
ntrn  <- 188-38 # reference 188*0.8
nCV <- 100 # times of CV

traits <- c("MajorAxis", "MinorAxis", "Perimeter")
# predictive correlation
corR1 <- matrix(0, ncol = ncol(y0), nrow = nCV) #OLS
  colnames(corR1) = traits

corR3 <- matrix(0, ncol = ncol(y0), nrow = nCV) #BayesC
  colnames(corR3) = traits

# CV
for (i in 1:nCV) {
  # random-sampling to decide testing & reference accessions
  set.seed(100 + i)
  index <- sample(1:nrow(y0), size=ntst) # random sampling
  y <- y0
  y[index,] <- NA
  for (j in 1:ncol(y0)){
  cat("Now running nCV = ", i,"trait = ", j, "\n")
  ySingle <- y[,j]
  
  #GBLUP
  fit1 <- BGLR(y=ySingle, ETA=ETA1, nIter = 30000,
             burnIn = 10000, thin = 5, verbose = F, saveAt = 'GBLUP_')
  pred1 = fit1$yHat
  corR1[i,j] <- cor(pred1[index], y0[index,j])
  
  
  #GMBLUP
  fit3 <- BGLR(y=ySingle, ETA=ETA3, nIter = 30000,
             burnIn = 10000, thin = 5, verbose = F, saveAt = 'GMBLUP_')
  pred3 = fit3$yHat
  corR3[i,j] <- cor(pred3[index], y0[index,j])
  }}

save(corR1, corR3, file=file.path(path.output, "stress_GBLUP_GMBLUP.rda"))
```


### 3.1.1 Calculate colmean
```{r}
load(file=file.path(path.output, "stress_GBLUP_GMBLUP.rda"))

colnames(corR1) = c("Major_Axis", "Minor_Axis", "Perimeter")
round(colMeans(corR1), digits = 4) #GBLUP

# colnames(corR2) = c("Major_Axis", "Minor_Axis", "Perimeter")
# round(colMeans(corR2), digits = 4)#MBLUP

colnames(corR3) = c("Major_Axis", "Minor_Axis", "Perimeter")
round(colMeans(corR3), digits = 4) #GMBLUP
```

### 3.1.2 Heritability
```{r, echo=TRUE, eval = FALSE}
# Scale traits (Normalize)
y0 = scale(trait.trt[,-c(1:2)], center = T, scale = T)

# G matrix
Gcs = scale(geno_trt, center = T, scale = T)
Gchnt = tcrossprod(Gcs)/ncol(Gcs)
EVD_Gchnt <- eigen(Gchnt)

# M matrix
# M matrix 
Mcs = scale(met.trt[,-c(1:2)]) #since met.rr was scaled when I did BLUP correction, so here no necessary to double normalization.
Mchnt = tcrossprod(Mcs) # for Bayesian MBLUP
Mchnt2 = Mchnt/mean(diag(Mchnt))
EVD_Mchnt <- eigen(Mchnt2)

#GBLUP
ETA1 <- list(
    G = list(V=EVD_Gchnt$vectors, d=EVD_Gchnt$values, model='RKHS')
   )

#MBLUP
ETA2 <- list(
        MET = list(V=EVD_Mchnt$vectors, d=EVD_Mchnt$values, model='RKHS')
   )

#GMBLUP
ETA3 <- list(
    G = list(V=EVD_Gchnt$vectors, d=EVD_Gchnt$values, model='RKHS'),
    MET = list(V=EVD_Mchnt$vectors, d=EVD_Mchnt$values, model='RKHS')
   )



for (j in 1:3){
  cat("Now running trait: ", colnames(y0)[j], "\n")
  set.seed(100+j)
  ySingle <- y0[,j]
  #GBLUP
  fit1 <- BGLR(y=ySingle, ETA=ETA1, nIter = 30000, 
             burnIn = 10000, thin = 5, verbose = F)
  
  varE = fit1$varE
  varM = fit1$ETA$G$varU
  h2M = varM/(varM+varE)
  cat("fit1 heritability for genomics is: ", round(h2M, 4), "\n")
  
  # MBLUP
  fit2 <- BGLR(y=ySingle, ETA=ETA2, nIter = 30000, 
             burnIn = 10000, thin = 5, verbose = F)
  varE = fit2$varE
  varM = fit2$ETA$MET$varU
  h2M = varM/(varM+varE)
  cat("fit2 heritability for metabolites is: ", round(h2M,4), "\n")
  
  #GMBLUP
  fit3 <- BGLR(y=ySingle, ETA=ETA3, nIter = 30000, 
             burnIn = 10000, thin = 5, verbose = F)
  varE = fit3$varE
  varM = fit3$ETA$MET$varU
  varG = fit3$ETA$G$varU
  h2G = varG/(varG+varM+varE)
  cat("fit3 heritability for genomics is: ", round(h2G,4), "\n")
  h2M= varM/(varG+varE+varM)
  cat("fit3 heritability for metabolites is: ", round(h2M,4), "\n")
  }
```


# 4. Draw plots
```{r}
######################################
###loading MBLUP and GBLUP GMBLUP#####
######################################

load(file=file.path(path.output, "control_OLS_MBLUP_BayesC.rda"))
corR_MBLUP_con = corR2
load(file=file.path(path.output, "control_GBLUP_GMBLUP.rda"))
corR_GBLUP_con = corR1
corR_GMBLUP_con = corR3

load(file=file.path(path.output, "stress_OLS_MBLUP_BayesC.rda"))
corR_MBLUP_trt = corR2
load(file=file.path(path.output, "stress_GBLUP_GMBLUP.rda"))
corR_GBLUP_trt = corR1
corR_GMBLUP_trt = corR3

###############################################
## Percentage Compare GBLUP, GMBLUP, MBLUP#####
###############################################

GBLUP <- rbind(cbind.data.frame(Treatment = "Control", corR_GBLUP_con),cbind.data.frame(Treatment = "Stress", corR_GBLUP_trt))
MBLUP <- rbind(cbind.data.frame(Treatment = "Control", corR_MBLUP_con),cbind.data.frame(Treatment = "Stress", corR_MBLUP_trt))
GMBLUP <- rbind(cbind.data.frame(Treatment = "Control", corR_GMBLUP_con),cbind.data.frame(Treatment = "Stress", corR_GMBLUP_trt))
traits = c("MajorAxis", "MinorAxis", "Perimeter")
# labels = c("  Grain length", "  Grain width", "Grain perimeter")
m=1
traitL <- list()
for( i in traits){

    BLUPcorR <- data.frame(GBLUP[,c("Treatment",i)], MBLUP[, i], GMBLUP[,i])
    BLUPcorR$Treatment <- as.factor(BLUPcorR$Treatment)
    colnames(BLUPcorR) <- c("Treatment", "GBLUP", "MBLUP", "GMBLUP")
    models <- c( "MBLUP", "GMBLUP", "GBLUP")
    comb_mol <- matrix(c("GBLUP", "MBLUP", "MBLUP", "GMBLUP", "GBLUP", "GMBLUP"), 2, 3)
    comb_mol_con <- rbind.data.frame(comb_mol, NA)
    comb_mol_trt <- rbind.data.frame(comb_mol, NA) #first row - second rw.


    BLUPcorR_con <- BLUPcorR %>% filter(Treatment == "Control")
    BLUPcorR_trt <- BLUPcorR %>% filter(Treatment == "Stress")

    for (j in 1:3){
      # j=1
      x=comb_mol[1,j]
      y=comb_mol[2,j]
      comb_mol_con[3,j] <- sum(BLUPcorR_con[,x]-BLUPcorR_con[,y]>0)
      comb_mol_trt[3,j] <- sum(BLUPcorR_trt[,x]-BLUPcorR_trt[,y]>0)

    }

    traitL[[m]] <- list(control = comb_mol_con, stress = comb_mol_trt)
    m = m+1

}

names(traitL) <- traits
GtraitL <- traitL
save(GtraitL, file="./compare_genomic_models_percentage.Rdata")


load("./compare_genomic_models_percentage.Rdata")

####################
## Draw plots#######
####################


GBLUP <- rbind(cbind.data.frame(Treatment = "Control", corR_GBLUP_con),cbind.data.frame(Treatment = "Stress", corR_GBLUP_trt))
MBLUP <- rbind(cbind.data.frame(Treatment = "Control", corR_MBLUP_con),cbind.data.frame(Treatment = "Stress", corR_MBLUP_trt))
GMBLUP <- rbind(cbind.data.frame(Treatment = "Control", corR_GMBLUP_con),cbind.data.frame(Treatment = "Stress", corR_GMBLUP_trt))
traits = c("MajorAxis", "MinorAxis", "Perimeter")
labels = c("  Grain length", "  Grain width", "Grain perimeter")

pp <- list()
t=1
for( i in traits){
    
    BLUPcorR <- data.frame(GBLUP[,c("Treatment",i)], MBLUP[, i], GMBLUP[,i])
    BLUPcorR$Treatment <- as.factor(BLUPcorR$Treatment)
    colnames(BLUPcorR) <- c("Treatment", "GBLUP", "MBLUP", "GMBLUP")
    models <- c( "MBLUP", "GMBLUP", "GBLUP")
    comb_mol <- matrix(c("GBLUP", "MBLUP", "MBLUP", "GMBLUP", "GBLUP", "GMBLUP"), 2, 3)
    
    temp = GtraitL[[i]]
    p=list()
    aa=array()

    for (j in 1:3){
      BLUPcorR$Treatment <- rep(c(paste0("Control: ", temp$control[3, j],"%"), paste0("HNT: ", temp$stress[3, j], "%")), each =100)
      p[[j]] <- ggplot(data = BLUPcorR, aes_string(comb_mol[1,j],comb_mol[2,j])) + 
                coord_fixed() +
                scale_x_continuous(limits=c(-0.2, 0.9), breaks=c(-0.2,0,0.2,0.4,0.6,0.8)) +
                scale_y_continuous(limits=c(-0.2, 0.9), breaks=c(-0.2,0,0.2,0.4,0.6,0.8)) +

                geom_point(size=2, alpha=0.5, aes(colour=Treatment)) + 
                geom_abline(intercept = 0, slope = 1)  + theme_bw() +
                labs(x=comb_mol[1,j],y=comb_mol[2,j])+
        
                theme(legend.position=c(0.815,0.08))+
                # scale_color_discrete(name="") +
                theme(legend.key.size = unit(0.01, "cm"), legend.title = element_blank())+
                theme(legend.background = element_rect(fill="aliceblue"))+
                theme(legend.text = element_text(size=10, face="bold"))
      # print(p[[j]])
    }
    
  # plot_Ave <- ggarrange(p[[1]],p[[2]],p[[3]], labels=i, nrow = 1, ncol = 3, common.legend = F)
  # print(plot_Ave)
  pp[[t]] <- ggarrange(p[[1]],p[[2]],p[[3]], labels=labels[t], nrow = 1, ncol = 3, common.legend = F) 
  t=t+1
# dev.print(pdf, file = file.path(path.output, paste0(i,"_GBLUP_MBLUP_GMBLUP_comb.pdf")), height = 5, width = 12)

 
}

plot1 <- ggarrange(pp[[1]],pp[[2]],pp[[3]], nrow = 3, ncol = 1)
print(plot1)
dev.print(pdf, file = file.path(path.output, "GBLUP_MBLUP_GMBLUP_comb.pdf"), height = 14, width = 12)
```

## 5. Montel test for M and G matrix
```{r eval=F}
library(vegan)
G1 = dist(Gchnt)
M1 = dist(Mchnt)
mantel(M1, M1, method="pearson", permutations=999, strata = NULL,
na.rm = FALSE)
```

# 6.Conventional H2
```{r}
library(lme4)
library(readxl)
Grain_Size <- read_excel("~/Library/CloudStorage/OneDrive-VirginiaTech/Research/Codes/research/RiceUNLMetabolites/Trait/Waseem_trait/Grain_Size_HNT_JS.xlsx", 
    sheet = "AllSeeds")
colnames(Grain_Size)[2] = "GSOR.ID"
Grain_Size$GSOR.ID = paste0("GSOR", Grain_Size$GSOR.ID)
Grain_size.con = Grain_Size[Grain_Size$Treatment == "C", ]
Grain_size.trt = Grain_Size[Grain_Size$Treatment == "S", ]

Grain_size.rep.con = Grain_size.con%>%group_by(GSOR.ID) %>% summarise(nn = length(unique(PlantNo.)))
rep.con = mean(Grain_size.rep.con$nn)
Grain_size.rep.trt = Grain_size.trt%>%group_by(GSOR.ID) %>% summarise(nn = length(unique(PlantNo.)))
rep.trt = mean(Grain_size.rep.trt$nn)

##### Control
Grain_size.con$GSOR.ID = as.factor(Grain_size.con$GSOR.ID)
Grain_size.con$PlantNo. = as.factor(Grain_size.con$PlantNo.)
traits = c("MajorAxisLength", "MinorAxisLength", "Perimeter")
for (i in 1:3){
formu = as.formula(paste0(traits[i],"~ 1 + (1|GSOR.ID) + PlantNo."))
fit.lme = lmer(formula = formu, data = Grain_size.con)
summary(fit.lme)
temp = as.data.frame(VarCorr(fit.lme))

H2 = temp[1,4]/(temp[1,4]+temp[2,4]/rep.con)
print(round(H2, 2))
}

##### Stress
Grain_size.trt$GSOR.ID = as.factor(Grain_size.trt$GSOR.ID)
Grain_size.trt$PlantNo. = as.factor(Grain_size.trt$PlantNo.)
traits = c("MajorAxisLength", "MinorAxisLength", "Perimeter")
for (i in 1:3){
formu = as.formula(paste0(traits[i],"~ 1 + (1|GSOR.ID) + PlantNo."))
fit.lme = lmer(formula = formu, data = Grain_size.trt)
summary(fit.lme)
temp = as.data.frame(VarCorr(fit.lme))

H2 = temp[1,4]/(temp[1,4]+temp[2,4]/rep.trt)
print(round(H2, 2))
}
```